---
title: "Urban Greenspace Thermal Ranking — Santiago"
subtitle: "All MINVU parks classified by vegetation structure and surface temperature (StgoHOT, summer 2024)"
author: "Urban Soil Biophysics Lab — CEDEUS / PUC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    page-layout: full
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(sf)
library(terra)
library(tidyverse)
library(exactextractr)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(scales)
library(patchwork)
library(gt)

options(exactextractr.progress = FALSE)
sf_use_s2(FALSE)
```

## Overview

This document maps all urban greenspaces in the Santiago Metropolitan Region
(MINVU registry) and ranks them by surface temperature extracted from the
**SantiagoHOT** dataset — a high-resolution (10 m) thermal model covering
three moments of a summer day in 2024: morning (AM · 07–08 h),
afternoon (AF · 16–17 h), and evening (PM · 19–20 h).

Parks are colour-coded by their dominant vegetation structure
(**woody** vs **herbaceous/grass**) and temperature decile within each group.
Priority sampling sites selected for field soil analysis are highlighted
as circle markers. Renders on committed files only — no GEE credentials needed.

> **Source code**: [Urban-Soil-Biophysics/stgohot-greenspaces](https://github.com/Urban-Soil-Biophysics/stgohot-greenspaces)

---

## 1 · Load and classify park polygons

```{r load-parks}
parks_raw <- st_read(
  "data/geo-data/parques_urbanos_minvu.geojson",
  quiet = TRUE
)

# Base R only — dplyr verbs cause list-coercion on this sf object
parks_r13 <- parks_raw[
  !is.na(parks_raw$REGION) & parks_raw$REGION == 13,
]

parks_r13$greenspace_type <- dplyr::case_when(
  stringr::str_detect(parks_r13$NOMBRE,
    regex("parque", ignore_case = TRUE)) ~ "Urban Park",
  stringr::str_detect(parks_r13$NOMBRE,
    regex("plaza|plazoleta|jardin|jardines|santuario|estadio|deporte|deportivo|campo|piscina",
          ignore_case = TRUE)) ~ "Lawn Park",
  stringr::str_detect(parks_r13$NOMBRE,
    regex("bandejon|bandejón|rotonda|canal|estero|autopista",
          ignore_case = TRUE)) ~ "Roadside",
  TRUE ~ NA_character_
)

parks_r13 <- parks_r13[!is.na(parks_r13$greenspace_type), ]
parks_r13$park_id <- seq_len(nrow(parks_r13))
parks_r13$area_m2 <- as.numeric(parks_r13$SUPERFICIE)
parks_r13$area_ha <- round(parks_r13$area_m2 / 1e4, 2)
parks_r13$name    <- as.character(parks_r13$NOMBRE)
parks_r13$commune <- as.character(parks_r13$COM)

# Keep only atomic columns — drops all list columns from source geojson
keep_cols <- c("park_id", "name", "commune",
               "area_m2", "area_ha", "greenspace_type")
parks_clean <- parks_r13[, keep_cols]

# Rebuild geometry cleanly through WKB
geom_wkb   <- sf::st_as_binary(sf::st_geometry(parks_clean))
geom_clean <- sf::st_as_sfc(geom_wkb, crs = 4326)
parks <- sf::st_set_geometry(
  sf::st_drop_geometry(parks_clean),
  geom_clean
) |> sf::st_as_sf()

sprintf("Parks loaded: %d | Types: %s",
        nrow(parks),
        paste(unique(parks$greenspace_type), collapse = ", "))
```

---

## 2 · Extract StgoHOT surface temperatures

```{r extract-temp}
stgo_am <- rast("data/stgo-hot/santiago-chile_am_temp_c.tif")
stgo_af <- rast("data/stgo-hot/santiago-chile_af_temp_c.tif")
stgo_pm <- rast("data/stgo-hot/santiago-chile_pm_temp_c.tif")

parks_utm <- sf::st_transform(parks, 32719)

temp_am <- exact_extract(stgo_am, parks_utm, "mean", progress = FALSE)
temp_af <- exact_extract(stgo_af, parks_utm, "mean", progress = FALSE)
temp_pm <- exact_extract(stgo_pm, parks_utm, "mean", progress = FALSE)

# Build as plain dataframe — avoid sf method dispatch issues
parks_thermal_df <- sf::st_drop_geometry(parks_utm)
parks_thermal_df$temp_am_c    <- temp_am
parks_thermal_df$temp_af_c    <- temp_af
parks_thermal_df$temp_pm_c    <- temp_pm
parks_thermal_df$temp_range_c <- temp_af - temp_am

has_temp   <- !is.na(parks_thermal_df$temp_af_c)
thermal_df <- parks_thermal_df[has_temp, ]
geom_therm <- geom_clean[has_temp]

parks_thermal <- sf::st_set_geometry(thermal_df, geom_therm) |>
  sf::st_as_sf() |>
  sf::st_transform(4326)

sprintf("Parks with temperature data: %d", nrow(parks_thermal))
```

---

## 3 · Vegetation structure

```{r vegetation}
ranking_path <- "data/results/unified_park_ranking.rds"
rds_path     <- "data/results/ugs_analysis_10m_utm.rds"
csv_path     <- "data/results/FIELD_SAMPLING_FINAL_REPORT.csv"

veg_cols <- c("mean_woody", "mean_grass", "woody_ratio", "grass_ratio",
              "structure_cat", "stgohot_decile_mean", "condition",
              "sampling_priority", "ndvi_mean", "ndvi_decile",
              "woody_decile", "temp_decile", "priority_score")

if (file.exists(ranking_path)) {
  veg <- readRDS(ranking_path) |>
    dplyr::select(park_id, dplyr::any_of(veg_cols))
} else if (file.exists(rds_path)) {
  veg <- readRDS(rds_path) |>
    sf::st_drop_geometry() |>
    dplyr::select(park_id, dplyr::any_of(veg_cols))
} else {
  veg <- tibble::tibble(park_id = thermal_df$park_id)
}

# Add missing columns as NA
for (col in veg_cols) {
  if (!col %in% names(veg)) veg[[col]] <- NA
}

# Plain dataframe join
full_df <- dplyr::left_join(thermal_df, veg, by = "park_id")

# Structural fallback
full_df$structure_cat <- dplyr::coalesce(
  full_df$structure_cat,
  dplyr::case_when(
    full_df$greenspace_type == "Lawn Park" ~ "High Herbaceous Cover",
    full_df$greenspace_type == "Roadside"  ~ "High Herbaceous Cover",
    TRUE                                   ~ NA_character_
  )
)

sprintf("Parks with structure data: %d | unclassified: %d",
        sum(!is.na(full_df$structure_cat)),
        sum( is.na(full_df$structure_cat)))
```

---

## 4 · Temperature ranking

```{r ranking}
csv_mgmt <- if (file.exists(csv_path)) {
  readr::read_csv(csv_path, show_col_types = FALSE) |>
    dplyr::select(park_id, management = Management)
} else {
  tibble::tibble(park_id = integer(), management = character())
}

ranked_df <- full_df |>
  dplyr::left_join(csv_mgmt, by = "park_id") |>
  dplyr::mutate(
    veg_group = dplyr::case_when(
      structure_cat == "High Woody Cover"      ~ "Woody",
      structure_cat == "High Herbaceous Cover" ~ "Grass",
      structure_cat == "Mixed/Balanced"        ~ "Mixed",
      TRUE                                     ~ "Unclassified"
    ),
    temp_decile_map = dplyr::if_else(
      veg_group != "Unclassified",
      dplyr::ntile(temp_af_c, 10),
      NA_integer_
    ),
    temp_label = dplyr::case_when(
      temp_decile_map %in% 1:3  ~ "Cool",
      temp_decile_map %in% 4:7  ~ "Moderate",
      temp_decile_map %in% 8:10 ~ "Hot",
      TRUE                      ~ NA_character_
    ),
    is_priority = !is.na(management)
  )

sprintf("ranked_df rows: %d", nrow(ranked_df))
```

---

## 5 · Interactive map

```{r palettes}
#| include: false
pal_woody <- colorNumeric(c("#c7e9b4","#41ab5d","#006d2c","#00441b"),
                          domain = 1:10, na.color = "#cccccc")
pal_grass <- colorNumeric(c("#fee5d9","#fc9272","#de2d26","#67000d"),
                          domain = 1:10, na.color = "#cccccc")
pal_mixed <- colorNumeric(c("#efedf5","#9e9ac8","#54278f"),
                          domain = 1:10, na.color = "#cccccc")

ranked_df <- ranked_df |>
  dplyr::mutate(
    fill_color = as.character(dplyr::case_when(
      veg_group == "Woody" ~ pal_woody(temp_decile_map),
      veg_group == "Grass" ~ pal_grass(temp_decile_map),
      veg_group == "Mixed" ~ pal_mixed(temp_decile_map),
      TRUE                 ~ "#cccccc"
    ))
  )
```

```{r popups}
#| include: false
make_popup <- function(name, commune, area_ha, type,
                       veg_group, structure_cat,
                       temp_am, temp_af, temp_pm, temp_range,
                       decile, temp_label,
                       woody_ratio, grass_ratio,
                       ndvi_mean, condition, sampling_priority,
                       is_priority, management) {

  badge <- if (isTRUE(is_priority))
    '<span style="background:#1a6b3a;color:#fff;padding:2px 7px;
     border-radius:10px;font-size:11px;font-weight:bold;">
     &#9733; Priority sampling site</span><br/>' else ""

  veg_row <- if (!is.na(woody_ratio))
    sprintf("<b>Woody:</b> %s &nbsp; <b>Grass:</b> %s<br/>",
            scales::percent(woody_ratio, accuracy = 1),
            scales::percent(grass_ratio, accuracy = 1)) else ""

  ndvi_row <- if (!is.na(ndvi_mean))
    sprintf("<b>NDVI:</b> %.3f<br/>", ndvi_mean) else ""

  temp_row <- if (!is.na(temp_af))
    sprintf("<b>AM:</b> %.1f &nbsp; <b>AF:</b> %.1f &nbsp; <b>PM:</b> %.1f °C<br/>
             <b>Range:</b> %.1f °C<br/>",
            temp_am, temp_af, temp_pm, temp_range) else "No temperature data"

  dcol <- dplyr::case_when(
    !is.na(temp_label) & temp_label == "Cool" ~ "#006d2c",
    !is.na(temp_label) & temp_label == "Hot"  ~ "#de2d26",
    TRUE ~ "#888"
  )
  dstr <- if (!is.na(decile)) sprintf("%d/10", as.integer(decile)) else "–"
  cond <- if (!is.na(condition))
    sprintf("<b>Condition:</b> %s<br/>", condition) else ""
  rank <- if (!is.na(sampling_priority))
    sprintf("<b>Priority:</b> %s<br/>", sampling_priority) else ""
  mgmt <- if (!is.na(management))
    sprintf("<b>Management:</b> %s<br/>", management) else ""

  sprintf(
    '<div style="font-family:sans-serif;font-size:13px;min-width:220px">
      %s<b style="font-size:14px">%s</b><br/>
      <span style="color:#777">%s &middot; %.2f ha &middot; %s</span>
      <hr style="margin:5px 0"/>
      <b>Structure:</b> %s <span style="color:#777">(%s)</span><br/>
      %s%s
      <hr style="margin:5px 0"/>%s
      <b>Temp decile (AF):</b> <span style="color:%s">%s %s</span><br/>
      %s%s%s
    </div>',
    badge, name, commune, area_ha, type,
    coalesce(structure_cat, "Unknown"), veg_group,
    veg_row, ndvi_row,
    temp_row, dcol, dstr, coalesce(temp_label, ""),
    cond, rank, mgmt
  )
}

ranked_df <- ranked_df |>
  dplyr::mutate(
    popup_html = purrr::pmap_chr(
      list(name, commune, area_ha, greenspace_type,
           veg_group, structure_cat,
           temp_am_c, temp_af_c, temp_pm_c, temp_range_c,
           temp_decile_map, temp_label,
           woody_ratio, grass_ratio,
           ndvi_mean, condition, sampling_priority,
           is_priority, management),
      make_popup
    )
  )
```

```{r leaflet-map}
# Centroid-based map — avoids expandLimitsBbox issues with MULTIPOLYGON
coords <- sf::st_coordinates(
  sf::st_centroid(sf::st_geometry(parks_thermal))
)

centroids <- ranked_df |>
  dplyr::mutate(lon = coords[, 1], lat = coords[, 2])

pal_map <- colorFactor(
  palette = c("#41ab5d", "#fc9272", "#9e9ac8", "#cccccc"),
  levels  = c("Woody", "Grass", "Mixed", "Unclassified")
)

leaflet(data = centroids) |>
  addProviderTiles("CartoDB.Positron",     group = "Light (CartoDB)") |>
  addProviderTiles("Esri.WorldImagery",    group = "Satellite (ESRI)") |>
  addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") |>
  addCircleMarkers(
    lng         = ~lon,
    lat         = ~lat,
    radius      = ~dplyr::if_else(is_priority, 10, 6),
    color       = ~dplyr::if_else(is_priority, "#fff", pal_map(veg_group)),
    weight      = ~dplyr::if_else(is_priority, 2, 0.5),
    fillColor   = ~pal_map(veg_group),
    fillOpacity = 0.8,
    popup       = ~popup_html,
    label       = ~name
  ) |>
  addLegend(
    position = "bottomleft",
    colors   = c("#41ab5d", "#fc9272", "#9e9ac8", "#cccccc", "#1a6b3a"),
    labels   = c("Woody", "Grass", "Mixed", "Unclassified", "★ Priority site"),
    title    = "Vegetation group",
    opacity  = 0.9
  ) |>
  addLayersControl(
    baseGroups = c("Light (CartoDB)", "Satellite (ESRI)", "OpenStreetMap"),
    options    = layersControlOptions(collapsed = FALSE)
  ) |>
  addScaleBar(position = "bottomright",
              options  = scaleBarOptions(imperial = FALSE)) |>
  setView(lng = -70.650, lat = -33.470, zoom = 11)
```

---

## 6 · Summary table

```{r summary-table}
ranked_df |>
  dplyr::filter(veg_group %in% c("Woody", "Grass")) |>
  dplyr::arrange(veg_group, temp_decile_map) |>
  dplyr::group_by(veg_group) |>
  dplyr::slice(c(1:5, (dplyr::n() - 4):dplyr::n())) |>
  dplyr::ungroup() |>
  dplyr::select(
    Group        = veg_group,
    Name         = name,
    Commune      = commune,
    `Area (ha)`  = area_ha,
    `AM (°C)`    = temp_am_c,
    `AF (°C)`    = temp_af_c,
    `PM (°C)`    = temp_pm_c,
    `Range (°C)` = temp_range_c,
    Decile       = temp_decile_map,
    Rank         = temp_label,
    Priority     = is_priority
  ) |>
  dplyr::mutate(
    dplyr::across(where(is.double), \(x) round(x, 1)),
    Priority = dplyr::if_else(Priority, "★", "")
  ) |>
  gt(groupname_col = "Group") |>
  tab_header(
    title    = "Extreme parks by temperature decile",
    subtitle = "Top 5 coolest and 5 hottest per vegetation group (AF)"
  ) |>
  cols_label(Priority = "★") |>
  tab_style(style     = cell_fill(color = "#e8f5e9"),
            locations = cells_body(rows = Rank == "Cool")) |>
  tab_style(style     = cell_fill(color = "#ffebee"),
            locations = cells_body(rows = Rank == "Hot")) |>
  tab_style(style     = cell_text(weight = "bold"),
            locations = cells_body(rows = Priority == "★")) |>
  tab_options(table.font.size = 13)
```

---

## 7 · Temperature distribution by vegetation group

```{r boxplot, fig.width=9, fig.height=5}
ranked_df |>
  dplyr::filter(veg_group %in% c("Woody", "Grass", "Mixed")) |>
  tidyr::pivot_longer(
    c(temp_am_c, temp_af_c, temp_pm_c),
    names_to = "moment", values_to = "temp_c"
  ) |>
  dplyr::mutate(
    moment = factor(moment,
      levels = c("temp_am_c", "temp_af_c", "temp_pm_c"),
      labels = c("Morning\n(07–08 h)",
                 "Afternoon\n(16–17 h)",
                 "Evening\n(19–20 h)"))
  ) |>
  ggplot(aes(x = veg_group, y = temp_c, fill = veg_group)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8, outlier.alpha = 0.4) +
  facet_wrap(~moment) +
  scale_fill_manual(
    values = c(Woody = "#41ab5d", Grass = "#fc9272", Mixed = "#9e9ac8")
  ) +
  labs(title    = "Surface temperature by vegetation structure",
       subtitle = "StgoHOT — all MINVU parks, Santiago RM, summer 2024",
       x = NULL, y = "Mean surface temperature (°C)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold"))
```

---

## 8 · Sampling priority ranking

```{r ranking-summary}
if (file.exists("data/results/summary_global.csv")) {
  readr::read_csv("data/results/summary_global.csv",
                  show_col_types = FALSE) |>
    gt() |>
    tab_header(title    = "Global sampling priority summary",
               subtitle = "By greenspace type and priority class") |>
    tab_options(table.font.size = 12)
} else {
  ranked_df |>
    dplyr::count(greenspace_type, temp_label) |>
    tidyr::drop_na() |>
    gt() |>
    tab_header(title    = "Park count by vegetation group and temperature class",
               subtitle = "Run 04_sample_sites.R for full priority ranking") |>
    tab_options(table.font.size = 12)
}
```

---

## 9 · NDVI — vegetation vigor

```{r ndvi-analysis, fig.width=10, fig.height=4}
ndvi_path <- "data/geo-data/NDVI_Santiago_Jan2024_30m.tif"

if (file.exists(ndvi_path)) {

  ndvi_rast     <- terra::rast(ndvi_path)
  parks_for_ndvi <- sf::st_transform(parks_thermal, 32719)
  ndvi_vals      <- exact_extract(ndvi_rast, parks_for_ndvi, "mean",
                                   progress = FALSE)

  ndvi_df <- ranked_df |>
    dplyr::mutate(ndvi_mean_calc = ndvi_vals) |>
    dplyr::filter(!is.na(ndvi_mean_calc), veg_group != "Unclassified")

  p_ndvi_type <- ndvi_df |>
    dplyr::filter(!is.na(greenspace_type)) |>
    ggplot(aes(x = greenspace_type, y = ndvi_mean_calc,
               fill = greenspace_type)) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.25, size = 1) +
    scale_fill_manual(values = c(
      "Urban Park" = "#2d6a4f",
      "Lawn Park"  = "#95d5b2",
      "Roadside"   = "#e76f51"
    )) +
    labs(title = "NDVI by park type",
         x = NULL, y = "Mean NDVI") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none")

  p_ndvi_veg <- ndvi_df |>
    ggplot(aes(x = veg_group, y = ndvi_mean_calc, fill = veg_group)) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.25, size = 1) +
    scale_fill_manual(
      values = c(Woody = "#41ab5d", Grass = "#fc9272", Mixed = "#9e9ac8")
    ) +
    labs(title = "NDVI by vegetation group",
         subtitle = "Landsat 8 — January 2024 (30m)",
         x = NULL, y = "Mean NDVI") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none")

  p_ndvi_temp <- ndvi_df |>
    dplyr::filter(!is.na(temp_af_c)) |>
    ggplot(aes(x = ndvi_mean_calc, y = temp_af_c, color = veg_group)) +
    geom_point(alpha = 0.4, size = 1.5) +
    geom_smooth(method = "lm", se = TRUE,
                color = "gray30", linewidth = 0.8) +
    scale_color_manual(
      values = c(Woody = "#41ab5d", Grass = "#fc9272", Mixed = "#9e9ac8")
    ) +
    labs(title = "NDVI vs afternoon temperature",
         x = "Mean NDVI", y = "AF temperature (°C)",
         color = NULL) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")

  p_ndvi_type + p_ndvi_veg + p_ndvi_temp +
    plot_layout(ncol = 3) +
    plot_annotation(
      title    = "NDVI — vegetation vigor across Santiago greenspaces",
      subtitle = "Landsat 8 & 9 Collection 2 Level 2 | January 2024 | 30m resolution"
    )

} else {
  cat("NDVI raster not found — expected at:", ndvi_path)
}
```

---

## 10 · Soil organic matter — field data association

```{r om-analysis, fig.width=10, fig.height=8}
om_joined_path  <- "data/results/om_ranking_joined.csv"
om_summary_path <- "data/results/om_ranking_summary.csv"

if (file.exists(om_joined_path)) {

  om_data <- readr::read_csv(om_joined_path, show_col_types = FALSE)

  p_type <- om_data |>
    dplyr::filter(!is.na(greenspace_type)) |>
    ggplot(aes(x = greenspace_type, y = om_value,
               fill = greenspace_type)) +
    geom_boxplot(alpha = 0.7, outlier.size = 1) +
    geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
    scale_fill_manual(values = c(
      "Urban Park" = "#2d6a4f",
      "Lawn Park"  = "#95d5b2",
      "Roadside"   = "#e76f51"
    )) +
    labs(title = "Soil OM by park type",
         x = NULL, y = "Organic matter (%)") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none")

  p_woody <- om_data |>
    dplyr::filter(!is.na(woody_ratio)) |>
    ggplot(aes(x = woody_ratio, y = om_value,
               color = greenspace_type)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_smooth(method = "lm", se = TRUE,
                color = "gray30", linewidth = 0.8) +
    scale_color_manual(values = c(
      "Urban Park" = "#2d6a4f",
      "Lawn Park"  = "#95d5b2",
      "Roadside"   = "#e76f51"
    )) +
    labs(title = "Soil OM ~ woody ratio",
         x = "Woody ratio (0–1)", y = "Organic matter (%)",
         color = NULL) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")

  p_temp <- om_data |>
    dplyr::filter(!is.na(stgohot_decile_mean)) |>
    ggplot(aes(x = stgohot_decile_mean, y = om_value,
               color = greenspace_type)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_smooth(method = "lm", se = TRUE,
                color = "gray30", linewidth = 0.8) +
    scale_x_continuous(breaks = 1:10) +
    scale_color_manual(values = c(
      "Urban Park" = "#2d6a4f",
      "Lawn Park"  = "#95d5b2",
      "Roadside"   = "#e76f51"
    )) +
    labs(title = "Soil OM ~ temperature decile",
         x = "StgoHOT mean decile (1=cool, 10=hot)",
         y = "Organic matter (%)", color = NULL) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")

  p_cond <- om_data |>
    dplyr::filter(!is.na(condition)) |>
    ggplot(aes(x = condition, y = om_value, fill = condition)) +
    geom_boxplot(alpha = 0.7, outlier.size = 1) +
    geom_jitter(width = 0.12, alpha = 0.4, size = 1.5) +
    scale_fill_manual(values = c(
      "Good condition" = "#52b788",
      "Poor condition" = "#e76f51"
    )) +
    labs(title = "Soil OM by park condition",
         x = NULL, y = "Organic matter (%)") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none")

  (p_type + p_cond) / (p_woody + p_temp) +
    plot_annotation(
      title    = "Soil organic matter associations with park characteristics",
      subtitle = sprintf(
        "Field data: n = %d points | Spatially joined to nearest MINVU park",
        nrow(om_data)
      )
    )

} else {
  cat("OM association file not found — run scripts/ranking/05_soil_om_association.R first.")
}
```

```{r om-summary-table}
if (file.exists(om_summary_path)) {
  readr::read_csv(om_summary_path, show_col_types = FALSE) |>
    gt() |>
    tab_header(
      title    = "Soil OM summary",
      subtitle = "Mean, SD and median OM (%) by park type and priority class"
    ) |>
    tab_options(table.font.size = 12)
}
```

---

## Reproducibility notes

| File | Source | Required |
|---|---|---|
| `data/geo-data/parques_urbanos_minvu.geojson` | Committed | ✅ |
| `data/stgo-hot/santiago-chile_{am,af,pm}_temp_c.tif` | Committed | ✅ |
| `data/results/FIELD_SAMPLING_FINAL_REPORT.csv` | Committed | ✅ |
| `data/geo-data/NDVI_Santiago_Jan2024_30m.tif` | Committed | Optional |
| `data/results/ugs_analysis_10m_utm.rds` | `03_zonal_statistics.R` | Optional |
| `data/results/intercept_stgohot_summary.rds` | `02_intercept-stgohot-ugs.R` | Optional |
| `data/results/unified_park_ranking.rds` | `04_sample_sites.R` | Optional |
| `data/results/om_ranking_joined.csv` | `05_soil_om_association.R` | Optional |

```r
quarto::quarto_render("index.qmd")
```
