---
title: "Urban Greenspace Analysis — Santiago RM"
subtitle: "Outputs and interpretation: thermal exposure, vegetation structure, and soil organic matter"
author: "Urban Soil Biophysics Lab — CEDEUS / PUC"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    page-layout: full
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(sf)
library(terra)
library(tidyverse)
library(exactextractr)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(scales)
library(patchwork)
library(gt)

options(exactextractr.progress = FALSE)
sf_use_s2(FALSE)

# ── Shared helpers ────────────────────────────────────────────────────────────

# Load and clean MINVU parks (base R to avoid list-coercion)
load_minvu <- function() {
  raw <- sf::st_read("data/geo-data/parques_urbanos_minvu.geojson", quiet = TRUE)
  r13 <- raw[!is.na(raw$REGION) & raw$REGION == 13, ]
  r13$greenspace_type <- dplyr::case_when(
    stringr::str_detect(r13$NOMBRE, regex("parque", ignore_case = TRUE)) ~
      "Urban Park",
    stringr::str_detect(r13$NOMBRE,
      regex("plaza|plazoleta|jardin|jardines|santuario|estadio|deporte|deportivo|campo|piscina",
            ignore_case = TRUE)) ~ "Lawn Park",
    stringr::str_detect(r13$NOMBRE,
      regex("bandejon|bandejón|rotonda|canal|estero|autopista",
            ignore_case = TRUE)) ~ "Roadside",
    TRUE ~ NA_character_
  )
  r13 <- r13[!is.na(r13$greenspace_type), ]
  r13$park_id  <- seq_len(nrow(r13))
  r13$area_m2  <- as.numeric(r13$SUPERFICIE)
  r13$area_ha  <- round(r13$area_m2 / 1e4, 2)
  r13$name     <- as.character(r13$NOMBRE)
  r13$commune  <- as.character(r13$COM)
  r13[, c("park_id","name","commune","area_m2","area_ha","greenspace_type")] |>
    sf::st_as_sf() |>
    sf::st_transform(4326)
}

# Rebuild geometry through WKB — removes list-type coordinate internals
clean_geom <- function(sf_obj) {
  wkb <- sf::st_as_binary(sf::st_geometry(sf_obj))
  sf::st_set_geometry(sf_obj, sf::st_as_sfc(wkb, crs = 4326))
}

# Raster → leaflet-ready SpatRaster projected to EPSG:3857
prep_rast_leaflet <- function(path) {
  r <- terra::rast(path)
  terra::project(r, "EPSG:3857")
}

# Colour palette helpers
pal_type <- colorFactor(
  palette = c("#2d6a4f", "#95d5b2", "#e76f51"),
  levels  = c("Urban Park", "Lawn Park", "Roadside")
)
pal_cond <- colorFactor(
  palette = c("#52b788", "#e76f51"),
  levels  = c("Good condition", "Poor condition"),
  na.color = "#cccccc"
)
pal_veg <- colorFactor(
  palette = c("#41ab5d", "#fc9272", "#9e9ac8", "#cccccc"),
  levels  = c("Woody", "Grass", "Mixed", "Unclassified")
)
pal_decile <- colorNumeric(
  palette  = "YlOrRd",
  domain   = 1:10,
  na.color = "transparent"
)
pal_ndvi <- colorNumeric(
  palette  = c("#ece7f2","#a6bddb","#2ca25f","#006d2c"),
  domain   = c(0, 0.8),
  na.color = "transparent"
)

basemap_tiles <- function(map) {
  map |>
    addProviderTiles("CartoDB.Positron",     group = "Light") |>
    addProviderTiles("Esri.WorldImagery",    group = "Satellite") |>
    addProviderTiles("OpenStreetMap.Mapnik", group = "Streets")
}
layer_ctrl <- function(map, overlays) {
  map |>
    addLayersControl(
      baseGroups    = c("Light","Satellite","Streets"),
      overlayGroups = overlays,
      options       = layersControlOptions(collapsed = TRUE)
    ) |>
    addScaleBar(position = "bottomright",
                options  = scaleBarOptions(imperial = FALSE))
}

# ── Load master data ──────────────────────────────────────────────────────────
parks     <- load_minvu() |> clean_geom()
ranking   <- readRDS("data/results/unified_park_ranking.rds")
stgohot   <- readRDS("data/results/intercept_stgohot_summary.rds") |>
               dplyr::mutate(time = as.character(time))

# Join ranking to parks geometry
# Join ranking to parks geometry
parks_full <- parks |>
  dplyr::left_join(
    ranking |> dplyr::select(park_id, mean_woody, mean_grass,
                              woody_ratio, grass_ratio, structure_cat,
                              stgohot_decile_mean, sampling_priority,
                              woody_decile, temp_decile, priority_score),
    by = "park_id"
  ) |>
  clean_geom()

# Join condition from areas_verdes
areas_verdes <- sf::st_read(
  "data/geo-data/sup_areas_verdes_stgo.geojson", quiet = TRUE
) |>
  sf::st_transform(4326) |>
  sf::st_make_valid() |>
  dplyr::select(estado_2, clase) |>
  dplyr::mutate(
    condition = dplyr::case_when(
      estado_2 == "BUEN" ~ "Good condition",
      estado_2 == "MAL"  ~ "Poor condition",
      TRUE               ~ NA_character_
    )
  )

parks_full <- parks_full |>
  sf::st_join(
    areas_verdes |> dplyr::select(condition),
    join    = sf::st_intersects,
    left    = TRUE,
    largest = TRUE
  ) |>
  clean_geom()

# Vegetation group
parks_full <- parks_full |>
  dplyr::mutate(
    veg_group = dplyr::case_when(
      structure_cat == "High Woody Cover"      ~ "Woody",
      structure_cat == "High Herbaceous Cover" ~ "Grass",
      structure_cat == "Mixed/Balanced"        ~ "Mixed",
      TRUE                                     ~ "Unclassified"
    )
  )

# Park centroids (clean)
centroids_df <- parks_full |>
  dplyr::mutate(
    lon = sf::st_coordinates(sf::st_centroid(sf::st_geometry(parks_full)))[,1],
    lat = sf::st_coordinates(sf::st_centroid(sf::st_geometry(parks_full)))[,2]
  ) |>
  sf::st_drop_geometry()

# Extract thermal per park for display
stgo_am_r <- terra::rast("data/stgo-hot/santiago-chile_am_temp_c.tif")
stgo_af_r <- terra::rast("data/stgo-hot/santiago-chile_af_temp_c.tif")
stgo_pm_r <- terra::rast("data/stgo-hot/santiago-chile_pm_temp_c.tif")

parks_utm <- sf::st_transform(parks_full, 32719)
centroids_df$temp_am <- exactextractr::exact_extract(stgo_am_r, parks_utm, "mean", progress=FALSE)
centroids_df$temp_af <- exactextractr::exact_extract(stgo_af_r, parks_utm, "mean", progress=FALSE)
centroids_df$temp_pm <- exactextractr::exact_extract(stgo_pm_r, parks_utm, "mean", progress=FALSE)
centroids_df$temp_mean <- rowMeans(
  centroids_df[, c("temp_am","temp_af","temp_pm")], na.rm = TRUE
)

# NDVI per park
ndvi_r   <- terra::rast("data/geo-data/NDVI_Santiago_Jan2024_30m.tif")
centroids_df$ndvi_mean <- exactextractr::exact_extract(ndvi_r, parks_utm, "mean", progress=FALSE)

# Landsat + MODIS decile per park
landsat_r <- terra::rast("data/geo-data/landsat_lst_deciles_jan2024.tif")
modis_r   <- terra::rast("data/geo-data/modis_lst_deciles_jan2024.tif")
centroids_df$landsat_decile <- exactextractr::exact_extract(landsat_r, parks_utm, "mean", progress=FALSE)
centroids_df$modis_decile   <- exactextractr::exact_extract(modis_r,   parks_utm, "mean", progress=FALSE)

# StgoHOT deciles per park per time (from intercept summary)
stgohot_wide <- stgohot |>
  dplyr::select(park_id, time, dominant_decile) |>
  tidyr::pivot_wider(names_from = time, values_from = dominant_decile,
                     names_prefix = "stgohot_") |>
  dplyr::rename_with(tolower)

centroids_df <- centroids_df |>
  dplyr::left_join(stgohot_wide, by = "park_id")
```

---

# Part 1 — Satellite thermal comparison

> Each data source characterizes thermal exposure differently. SantiagoHOT provides air temperature at three moments of the day at high resolution (~10 m). Landsat 8 provides land surface temperature (LST) at 30 m. MODIS provides LST at 1 km — coarse enough that small parks inherit the thermal signature of surrounding streets.

## 1.1 SantiagoHOT — air temperature (AM · AF · PM)

```{r map-stgohot-rasters}
am_r  <- prep_rast_leaflet("data/stgo-hot/santiago-chile_am_temp_c.tif")
af_r  <- prep_rast_leaflet("data/stgo-hot/santiago-chile_af_temp_c.tif")
pm_r  <- prep_rast_leaflet("data/stgo-hot/santiago-chile_pm_temp_c.tif")

pal_temp <- colorNumeric("YlOrRd", values(am_r), na.color = "transparent")

leaflet() |>
  basemap_tiles() |>
  addRasterImage(am_r, colors = pal_temp, opacity = 0.7,
                 group = "AM (07–08 h)") |>
  addRasterImage(af_r, colors = pal_temp, opacity = 0.7,
                 group = "AF (15–16 h)") |>
  addRasterImage(pm_r, colors = pal_temp, opacity = 0.7,
                 group = "PM (19–20 h)") |>
  addLegend(pal = pal_temp, values = values(am_r),
            title = "Air temp (°C)", position = "bottomleft") |>
  layer_ctrl(c("AM (07–08 h)","AF (15–16 h)","PM (19–20 h)")) |>
  setView(-70.65, -33.47, zoom = 11)
```

> Temperature peaks at AF (afternoon). The city centre and eastern piedmont sectors consistently show higher temperatures across all three moments.

## 1.2 Landsat 8 — LST deciles (30 m)

```{r map-landsat}
ls_r  <- prep_rast_leaflet("data/geo-data/landsat_lst_deciles_jan2024.tif")

leaflet() |>
  basemap_tiles() |>
  addRasterImage(ls_r, colors = pal_decile, opacity = 0.7,
                 group = "Landsat LST deciles") |>
  addLegend(pal = pal_decile, values = 1:10,
            title = "LST decile<br/>(1=cool, 10=hot)",
            position = "bottomleft") |>
  layer_ctrl("Landsat LST deciles") |>
  setView(-70.65, -33.47, zoom = 11)
```

> At 30 m, Landsat resolves within-park variability. Deciles are computed from the least-cloudy January 2024 image.

## 1.3 MODIS — LST deciles (1 km)

```{r map-modis}
mo_r  <- prep_rast_leaflet("data/geo-data/modis_lst_deciles_jan2024.tif")

leaflet() |>
  basemap_tiles() |>
  addRasterImage(mo_r, colors = pal_decile, opacity = 0.7,
                 group = "MODIS LST deciles") |>
  addLegend(pal = pal_decile, values = 1:10,
            title = "LST decile<br/>(1=cool, 10=hot)",
            position = "bottomleft") |>
  layer_ctrl("MODIS LST deciles") |>
  setView(-70.65, -33.47, zoom = 11)
```

> MODIS pixels are 1 km² — most plazas and roadsides fall within a single pixel shared with streets, inflating their apparent LST. Parks < 1 ha should be interpreted with caution.

---

# Part 2 — NDVI — vegetation vigor

```{r map-ndvi}
nd_r <- prep_rast_leaflet("data/geo-data/NDVI_Santiago_Jan2024_30m.tif")

leaflet() |>
  basemap_tiles() |>
  addRasterImage(nd_r, colors = pal_ndvi, opacity = 0.8,
                 group = "NDVI") |>
  addLegend(pal = pal_ndvi, values = c(0, 0.8),
            title = "NDVI",
            labFormat = labelFormat(digits = 2),
            position = "bottomleft") |>
  layer_ctrl("NDVI") |>
  setView(-70.65, -33.47, zoom = 11)
```

> NDVI from Landsat 8 & 9 (January 2024, 30 m). Values near 0.2 indicate impervious or bare soil; values above 0.5 indicate active canopy. Dense tree cover in Parque Metropolitano (NE) and irrigated parks in Las Condes stand out clearly.

---

# Part 3 — Park inventory

> MINVU registry for Region Metropolitana (updated 12/07/2022). Parks classified into Urban Parks, Lawn Parks, and Roadsides by name pattern matching. Condition (Good / Poor) joined from the áreas verdes dataset via spatial intersection.

## 3.1 Park type

```{r map-park-type}
parks_type <- parks_full |>
  dplyr::select(park_id, name, commune, area_ha, greenspace_type) |>
  clean_geom()

leaflet(parks_type) |>
  basemap_tiles() |>
  addPolygons(
    fillColor   = ~pal_type(greenspace_type),
    fillOpacity = 0.7,
    color       = "white", weight = 0.3,
    popup = ~sprintf(
      "<b>%s</b><br/>%s<br/>%.2f ha<br/><i>%s</i>",
      name, commune, area_ha, greenspace_type
    ),
    label = ~name,
    group = "Park type"
  ) |>
  addLegend(pal = pal_type, values = ~greenspace_type,
            title = "Park type", position = "bottomleft") |>
  layer_ctrl("Park type") |>
  setView(-70.65, -33.47, zoom = 11)
```

## 3.2 Park condition (Good / Poor)

```{r map-park-condition}
parks_cond <- parks_full |>
  dplyr::select(park_id, name, commune, area_ha, condition) |>
  clean_geom()

leaflet(parks_cond) |>
  basemap_tiles() |>
  addPolygons(
    fillColor   = ~pal_cond(condition),
    fillOpacity = 0.7,
    color       = "white", weight = 0.3,
    popup = ~sprintf(
      "<b>%s</b><br/>%s<br/>%.2f ha<br/><i>%s</i>",
      name, commune, area_ha, coalesce(condition, "No data")
    ),
    label = ~name,
    group = "Condition"
  ) |>
  addLegend(pal = pal_cond, values = ~condition,
            title = "Condition", position = "bottomleft",
            na.label = "No data") |>
  layer_ctrl("Condition") |>
  setView(-70.65, -33.47, zoom = 11)
```

> 75 % of parks with condition data are classified as Good (BUEN). Deteriorated parks cluster in lower-income comunas in the south and west.

---

# Part 4 — Dynamic World vegetation structure

> Woody ratio = (trees + shrubs) / (trees + shrubs + grass) from Google Dynamic World at 10 m resolution. Values above 0.6 = High Woody Cover; below 0.4 = High Herbaceous Cover.

## 4.1 Vegetation group per park

```{r map-veg-group}
parks_veg <- parks_full |>
  dplyr::select(park_id, name, commune, area_ha,
                veg_group, woody_ratio, grass_ratio) |>
  dplyr::mutate(
    woody_pct = scales::percent(woody_ratio, accuracy = 1),
    grass_pct = scales::percent(grass_ratio, accuracy = 1)
  ) |>
  clean_geom()

leaflet(parks_veg) |>
  basemap_tiles() |>
  addPolygons(
    fillColor   = ~pal_veg(veg_group),
    fillOpacity = 0.75,
    color       = "white", weight = 0.3,
    popup = ~sprintf(
      "<b>%s</b><br/>%s · %.2f ha<br/>
       <b>Woody:</b> %s &nbsp; <b>Grass:</b> %s<br/>
       <i>%s</i>",
      name, commune, area_ha,
      woody_pct, grass_pct, veg_group
    ),
    label = ~name,
    group = "Vegetation group"
  ) |>
  addLegend(pal = pal_veg, values = ~veg_group,
            title = "Vegetation group", position = "bottomleft") |>
  layer_ctrl("Vegetation group") |>
  setView(-70.65, -33.47, zoom = 11)
```


> Most Urban Parks fall in the Woody category. The majority of Lawn Parks and Roadsides are Grass or Unclassified, consistent with their management as mowed turf or paved medians.

---

# Part 5 — Combined thermal leaflet

> All thermal sources on a single map. Toggle between layers to compare how Landsat, MODIS, and SantiagoHOT characterise the same parks.

```{r map-thermal-combined}
pal_am  <- colorNumeric("Blues",  domain = range(centroids_df$temp_am,  na.rm=TRUE), na.color="transparent")
pal_af  <- colorNumeric("YlOrRd", domain = range(centroids_df$temp_af,  na.rm=TRUE), na.color="transparent")
pal_pm  <- colorNumeric("Purples",domain = range(centroids_df$temp_pm,  na.rm=TRUE), na.color="transparent")
pal_avg <- colorNumeric("RdYlBu", domain = range(centroids_df$temp_mean,na.rm=TRUE),
                        reverse = TRUE, na.color="transparent")
pal_ls  <- colorNumeric("YlOrRd", domain = 1:10, na.color="transparent")
pal_mo  <- colorNumeric("YlOrRd", domain = 1:10, na.color="transparent")

leaflet(centroids_df) |>
  basemap_tiles() |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_am(temp_am), fillColor = ~pal_am(temp_am),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>AM: %.1f °C", name, temp_am),
    group = "StgoHOT AM") |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_af(temp_af), fillColor = ~pal_af(temp_af),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>AF: %.1f °C", name, temp_af),
    group = "StgoHOT AF") |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_pm(temp_pm), fillColor = ~pal_pm(temp_pm),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>PM: %.1f °C", name, temp_pm),
    group = "StgoHOT PM") |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_avg(temp_mean), fillColor = ~pal_avg(temp_mean),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>Mean: %.1f °C", name, temp_mean),
    group = "StgoHOT Mean") |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_ls(landsat_decile), fillColor = ~pal_ls(landsat_decile),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>Landsat decile: %.1f", name, landsat_decile),
    group = "Landsat decile") |>
  addCircleMarkers(~lon, ~lat, radius = 5,
    color = ~pal_mo(modis_decile), fillColor = ~pal_mo(modis_decile),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~sprintf("<b>%s</b><br/>MODIS decile: %.1f", name, modis_decile),
    group = "MODIS decile") |>
  addLayersControl(
    baseGroups    = c("Light","Satellite","Streets"),
    overlayGroups = c("StgoHOT AM","StgoHOT AF","StgoHOT PM",
                      "StgoHOT Mean","Landsat decile","MODIS decile"),
    options       = layersControlOptions(collapsed = FALSE)
  ) |>
  addScaleBar(position = "bottomright",
              options  = scaleBarOptions(imperial = FALSE)) |>
  setView(-70.65, -33.47, zoom = 11)
```

> Agreement between sources is generally high for large parks (> 1 ha). For small plazas and roadsides, MODIS deciles diverge from SantiagoHOT due to the 1 km pixel size.

---

# Part 6 — Park characteristics combined

> Integrated view of park type, management condition, NDVI, and vegetation structure (woody/grass/mixed). Click any marker for details.

```{r map-park-chars}
pal_ndvi_pts <- colorNumeric(c("#ece7f2","#41ab5d","#006d2c"),
                              domain = c(0,0.8), na.color="#cccccc")

centroids_df <- centroids_df |>
  dplyr::mutate(
    popup_chars = sprintf(
      '<b>%s</b><br/>%s · %.2f ha<br/>
       <b>Type:</b> %s<br/>
       <b>Condition:</b> %s<br/>
       <b>Veg group:</b> %s<br/>
       <b>NDVI:</b> %.3f<br/>
       <b>Woody ratio:</b> %s',
      name, commune, area_ha,
      greenspace_type,
      coalesce(condition, "No data"),
      veg_group,
      coalesce(ndvi_mean, NA_real_),
      scales::percent(coalesce(woody_ratio, NA_real_), accuracy=1)
    )
  )

leaflet(centroids_df) |>
  basemap_tiles() |>
  addCircleMarkers(~lon, ~lat, radius = 6,
    color = ~pal_type(greenspace_type),
    fillColor = ~pal_type(greenspace_type),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~popup_chars, label = ~name,
    group = "Park type") |>
  addCircleMarkers(~lon, ~lat, radius = 6,
    color = ~pal_cond(condition),
    fillColor = ~pal_cond(condition),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~popup_chars, label = ~name,
    group = "Condition") |>
  addCircleMarkers(~lon, ~lat, radius = 6,
    color = ~pal_ndvi_pts(ndvi_mean),
    fillColor = ~pal_ndvi_pts(ndvi_mean),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~popup_chars, label = ~name,
    group = "NDVI") |>
  addCircleMarkers(~lon, ~lat, radius = 6,
    color = ~pal_veg(veg_group),
    fillColor = ~pal_veg(veg_group),
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~popup_chars, label = ~name,
    group = "Veg group") |>
  addLayersControl(
    baseGroups    = c("Light","Satellite","Streets"),
    overlayGroups = c("Park type","Condition","NDVI","Veg group"),
    options       = layersControlOptions(collapsed = FALSE)
  ) |>
  addScaleBar(position = "bottomright",
              options  = scaleBarOptions(imperial = FALSE)) |>
  setView(-70.65, -33.47, zoom = 11)
```

---

# Part 7 — NDVI associations

## 7.1 NDVI by park type, condition, and vegetation group

```{r ndvi-boxplots, fig.width=11, fig.height=4}
ndvi_df <- centroids_df |>
  dplyr::filter(!is.na(ndvi_mean))

p_type <- ndvi_df |>
  dplyr::filter(!is.na(greenspace_type)) |>
  ggplot(aes(x = greenspace_type, y = ndvi_mean,
             fill = greenspace_type)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
  geom_jitter(width = 0.15, alpha = 0.2, size = 0.8) +
  scale_fill_manual(values = c(
    "Urban Park" = "#2d6a4f",
    "Lawn Park"  = "#95d5b2",
    "Roadside"   = "#e76f51"
  )) +
  labs(x = NULL, y = "Mean NDVI") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 20, hjust = 1))

p_cond <- ndvi_df |>
  dplyr::filter(!is.na(condition)) |>
  ggplot(aes(x = condition, y = ndvi_mean, fill = condition)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
  geom_jitter(width = 0.12, alpha = 0.2, size = 0.8) +
  scale_fill_manual(values = c(
    "Good condition" = "#52b788",
    "Poor condition" = "#e76f51"
  )) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 20, hjust = 1))

p_veg <- ndvi_df |>
  dplyr::filter(veg_group != "Unclassified") |>
  ggplot(aes(x = veg_group, y = ndvi_mean, fill = veg_group)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
  scale_fill_manual(values = c(
    "Woody" = "#41ab5d", "Grass" = "#fc9272", "Mixed" = "#9e9ac8"
  )) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

p_type + p_cond + p_veg +
  plot_annotation(
    title    = "NDVI by park characteristics",
    subtitle = "Landsat 8 & 9 — January 2024 (30 m)"
  )
```

> Urban Parks have the highest median NDVI, consistent with larger canopy cover. Good condition parks score slightly higher than poor condition ones. Woody parks unsurprisingly lead in vegetation vigor.

## 7.2 NDVI vs thermal sources

```{r ndvi-vs-thermal, fig.width=12, fig.height=4}
vars <- list(
  list(col="temp_af",       lab="AF temperature (°C)",    pal="YlOrRd"),
  list(col="landsat_decile",lab="Landsat decile",          pal="YlOrRd"),
  list(col="modis_decile",  lab="MODIS decile",            pal="YlOrRd")
)

make_scatter <- function(xcol, xlab) {
  ndvi_df |>
    dplyr::filter(!is.na(.data[[xcol]]), veg_group != "Unclassified") |>
    ggplot(aes(x = .data[[xcol]], y = ndvi_mean, color = veg_group)) +
    geom_point(alpha = 0.4, size = 1.2) +
    geom_smooth(method = "lm", se = TRUE,
                color = "gray30", linewidth = 0.7) +
    scale_color_manual(values = c(
      "Woody"="#41ab5d","Grass"="#fc9272",
      "Mixed"="#9e9ac8","Unclassified"="#cccccc"
    )) +
    labs(x = xlab, y = "Mean NDVI", color = NULL) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "bottom",
          legend.key.size  = unit(0.4,"cm"))
}

make_scatter("temp_af",        "AF temperature (°C)") +
make_scatter("landsat_decile", "Landsat LST decile")  +
make_scatter("modis_decile",   "MODIS LST decile")    +
  plot_layout(ncol = 3) +
  plot_annotation(
    title    = "NDVI vs thermal exposure",
    subtitle = "Negative slope expected: hotter parks tend to have lower vegetation vigor"
  )
```

---

# Part 8 — Field soil organic matter

> 59 soil samples from Santiago RM, spatially joined to the nearest MINVU park (nearest-feature, WGS84). Organic matter (OM %) measured in the field.

```{r load-om}
#| include: false
fix_coord <- function(x) {
  x_nodots <- stringr::str_remove_all(as.character(x), "\\.")
  as.numeric(sub("^(-?)(\\d{2})(\\d+)$", "\\1\\2.\\3", x_nodots))
}

om_path <- "data/geo-data/dataset_rm_test.csv"

if (file.exists(om_path)) {
  om_raw <- readr::read_csv(om_path, show_col_types = FALSE)
  om_clean <- om_raw |>
    dplyr::mutate(
      lat      = fix_coord(lat),
      lon      = fix_coord(lon),
      om_value = suppressWarnings(as.numeric(om))
    ) |>
    dplyr::filter(!is.na(lat), !is.na(lon), !is.na(om_value))

  om_sf <- sf::st_as_sf(om_clean, coords = c("lon","lat"), crs = 4326)

  # Build park centroids for join
  minvu_raw   <- sf::st_read("data/geo-data/parques_urbanos_minvu.geojson",
                              quiet = TRUE)
  r13         <- minvu_raw[!is.na(minvu_raw$REGION) & minvu_raw$REGION == 13, ]
  r13$name_r  <- as.character(r13$NOMBRE)
  ctrs        <- sf::st_centroid(r13)
  ctrs_df     <- ctrs |>
    dplyr::mutate(
      name_upper = toupper(name_r),
      lon_park   = sf::st_coordinates(ctrs)[,1],
      lat_park   = sf::st_coordinates(ctrs)[,2]
    ) |>
    sf::st_drop_geometry() |>
    dplyr::distinct(name_upper, .keep_all = TRUE)

# In the load-om chunk, replace the ranking_sf construction with:
ranking_sf <- ranking |>
  dplyr::mutate(name_upper = toupper(name)) |>
  dplyr::left_join(
    ctrs_df |> dplyr::select(name_upper, lon_park, lat_park),
    by = "name_upper", relationship = "many-to-one"
  ) |>
  dplyr::filter(!is.na(lon_park)) |>
  sf::st_as_sf(coords = c("lon_park","lat_park"), crs = 4326)

  om_joined <- sf::st_join(om_sf, ranking_sf, join = sf::st_nearest_feature) |>
    sf::st_drop_geometry() |>
    dplyr::bind_cols(
      sf::st_coordinates(om_sf) |>
        as.data.frame() |>
        dplyr::rename(lon = X, lat = Y)
    ) |>
    dplyr::mutate(
      veg_group = dplyr::case_when(
        structure_cat == "High Woody Cover"      ~ "Woody",
        structure_cat == "High Herbaceous Cover" ~ "Grass",
        structure_cat == "Mixed/Balanced"        ~ "Mixed",
        TRUE                                     ~ "Unclassified"
      )
    )
om_joined <- om_joined |>
  dplyr::left_join(
    centroids_df |>
      dplyr::select(park_id, ndvi_mean,
                    temp_mean, temp_am, temp_af, temp_pm,
                    landsat_decile, modis_decile),
    by = "park_id"
  )
  om_has_data <- TRUE
} else {
  om_has_data <- FALSE
}
```

## 8.1 Map of OM sampling sites

```{r map-om}
if (om_has_data) {
  pal_om <- colorNumeric("YlGn", domain = om_joined$om_value, na.color="#cccccc")

  leaflet(om_joined) |>
    basemap_tiles() |>
    addCircleMarkers(~lon, ~lat,
      radius      = ~scales::rescale(om_value, to = c(4,12)),
      fillColor   = ~pal_om(om_value),
      fillOpacity = 0.85,
      color       = "white", weight = 1,
      popup = ~sprintf(
        "<b>OM: %.2f %%</b><br/>
         <b>Park:</b> %s<br/>
         <b>Type:</b> %s<br/>
         <b>Condition:</b> %s<br/>
         <b>Veg group:</b> %s",
        om_value,
        coalesce(name, "—"),
        coalesce(greenspace_type, "—"),
        coalesce(condition, "No data"),
        veg_group
      ),
      group = "OM samples"
    ) |>
    addLegend(pal = pal_om, values = ~om_value,
              title = "Organic matter (%)",
              position = "bottomleft") |>
    layer_ctrl("OM samples") |>
    setView(-70.65, -33.47, zoom = 11)
} else {
  cat("dataset_rm_test.csv not found — add to data/geo-data/")
}
```

> Marker size proportional to OM value. Higher OM tends to cluster in wooded parks in the eastern sector (Providencia, Las Condes), consistent with higher canopy input and lower soil temperature.

## 8.2 OM associations

```{r om-plots, fig.width=12, fig.height=9}
if (om_has_data) {
  col_vals <- c("Urban Park"="#2d6a4f","Lawn Park"="#95d5b2","Roadside"="#e76f51")

  p1 <- om_joined |> dplyr::filter(!is.na(greenspace_type)) |>
    ggplot(aes(x=greenspace_type, y=om_value, fill=greenspace_type)) +
    geom_boxplot(alpha=0.7, outlier.size=1) +
    geom_jitter(width=0.15, alpha=0.5, size=1.5) +
    scale_fill_manual(values=col_vals) +
    labs(title="OM by park type", x=NULL, y="OM (%)") +
    theme_minimal(base_size=10) + theme(legend.position="none",
    axis.text.x=element_text(angle=20,hjust=1))

  p2 <- om_joined |> dplyr::filter(!is.na(condition)) |>
    ggplot(aes(x=condition, y=om_value, fill=condition)) +
    geom_boxplot(alpha=0.7, outlier.size=1) +
    geom_jitter(width=0.12, alpha=0.5, size=1.5) +
    scale_fill_manual(values=c("Good condition"="#52b788","Poor condition"="#e76f51")) +
    labs(title="OM by condition", x=NULL, y=NULL) +
    theme_minimal(base_size=10) + theme(legend.position="none",
    axis.text.x=element_text(angle=20,hjust=1))

  p3 <- om_joined |> dplyr::filter(veg_group!="Unclassified") |>
    ggplot(aes(x=veg_group, y=om_value, fill=veg_group)) +
    geom_boxplot(alpha=0.7, outlier.size=1) +
    scale_fill_manual(values=c("Woody"="#41ab5d","Grass"="#fc9272","Mixed"="#9e9ac8")) +
    labs(title="OM by veg group", x=NULL, y=NULL) +
    theme_minimal(base_size=10) + theme(legend.position="none")

  make_om_scatter <- function(xcol, xlab) {
    om_joined |>
      dplyr::filter(!is.na(.data[[xcol]]), !is.na(greenspace_type)) |>
      ggplot(aes(x=.data[[xcol]], y=om_value, color=greenspace_type)) +
      geom_point(alpha=0.7, size=2) +
      geom_smooth(method="lm", se=TRUE, color="gray30", linewidth=0.7) +
      scale_color_manual(values=col_vals) +
      labs(x=xlab, y="OM (%)", color=NULL) +
      theme_minimal(base_size=10) +
      theme(legend.position="bottom", legend.key.size=unit(0.4,"cm"))
  }

  p4 <- make_om_scatter("ndvi_mean",      "NDVI")
  p5 <- make_om_scatter("temp_mean",      "StgoHOT mean (°C)")
  p6 <- make_om_scatter("landsat_decile", "Landsat decile")
  p7 <- make_om_scatter("modis_decile",   "MODIS decile")

  (p1 | p2 | p3) /
  (p4 | p5 | p6) +
  plot_annotation(
    title    = "Soil organic matter associations",
    subtitle = sprintf("n = %d field samples | January 2024", nrow(om_joined))
  )
} else {
  cat("OM data not available.")
}
```

> Higher NDVI is the strongest predictor of OM in this preliminary dataset, consistent with the expected soil C input from leaf litter. Thermal exposure (StgoHOT, Landsat) shows a negative trend — hotter parks tend to have lower OM, likely through accelerated decomposition.

---

*Data: SantiagoHOT (Jan 20, 2024) · Landsat 8 C2 T1 L2 (Jan 2024) · MODIS MOD11A1 (Jan 20, 2024) · Google Dynamic World v1 (Jan 2024 ±5 d) · MINVU Urban Parks (updated 12/07/2022) · Áreas Verdes Santiago · Field sampling 2024.*
